---
title: "learning dplyr"
output: html_notebook
---

# Intro 

In this R tutorial we learn about dplyr and ggplot

The data set is originally from here: https://packages.revolutionanalytics.com/datasets/

First we load the necessary libraries:
```{r, echo=FALSE}
library(tidyverse)
```

Load the data
```{r}
load("flights.rda")
```
alternatively you can also just load the data by double clicking it.

Generate a summary
```{r}
summary(flights)
```

# select

Find distinct years
```{r}
flights %>% select(YEAR) %>% distinct()
```

Find unique months 
```{r}
flights %>% select(MONTH) %>% distinct()
```

Find all distinct carriers
```{r}
flights %>% select(UNIQUE_CARRIER) %>% distinct()
```

Find all columns containing the word delay and only show first five rows
```{r}
flights %>% select(contains("DELAY")) %>% head(5)
```

You can learn more about these helper functions such as "contains" in the cheatsheet
https://www.r-bloggers.com/the-complete-catalog-of-argument-variations-of-select-in-dplyr/

# filter

filter out only the flights of American Airlines
```{r}
flights %>% filter(UNIQUE_CARRIER == "AA") %>% head(5)
```

filter out only the flights of American Airlines in February
```{r}
flights %>% filter(UNIQUE_CARRIER == "AA" & MONTH == 2) %>% head(5)
```

filter flights that of American Airline that have not started from JFK
```{r}
flights %>% filter(UNIQUE_CARRIER == "AA" & ORIGIN != JFK) %>% head(5)
```

filter flights that have covered a distance of at least 300
```{r}
flights %>% filter(DISTANCE >= 300) %>% head(5)
```

# arrange

sort flights according to the day of month in ascending order
```{r}
flights %>% arrange(DAY_OF_MONTH) %>% head(5)
```

sort flights according to the day of month and month in ascending order
```{r}
flights %>% arrange(DAY_OF_MONTH,MONTH) %>% head(5)
```

sort flights according to month in descending order
```{r}
flights %>% arrange(desc(MONTH)) %>% head(5)
```

# mutate

Mutations allow us to define new variables

Define a variable delay_per_distance which is defined as the arrival delay over the distance travelled.
```{r}
flights %>% mutate(delay_per_distance = ARR_DELAY/DISTANCE) %>% head(5)
```
# summarise

compute the mean distance and standard deviation of all flights
```{r}
flights %>% summarise(mean_distance = mean(DISTANCE),sd_distance = sd(DISTANCE))
```

# grouping

Now we learn about grouping. The command group_by(data,A,B) groups data according to the values of A and B. Say A takes values in {0,1} and B={0,1,2} then group_by sets groups according to all possible combinations of these values. If we run a summarise command on a grouped data set we get a statistics for all samples in each group.

Lets compute the average distance per carrier
```{r}
flights %>% group_by(UNIQUE_CARRIER) %>% summarise(mean_distance_per_carrier=mean(DISTANCE))
```

Grouping allows us to run quite complex computations with very short code.

Find the carrier with largest average delay per distance, i.e. delay/distance
```{r}
flights %>% mutate(delay_by_distance = ARR_DELAY/DISTANCE) %>%
  group_by(UNIQUE_CARRIER) %>%
  summarise(average_del_by_dist = mean(delay_by_distance)) %>%
  arrange(desc(average_del_by_dist))
```
